---
title: "Reproducible documents"
author: "Anne G Jensen"
format: html
---

## Basics of R

```{r}
1 + 1
```

## Setup code chunk
```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |> rename(auc_glucose = auc_pg, 
                                                        auc_insulin = auc_ins) |>
  filter(glykemi == 0) |> 
  select(-glykemi)
```



## Showing the data raw 

```{r first option}
post_meal_data
```

```{r glimpse option}
glimpse(post_meal_data)
```


?colnames


## Bold and kursiv 
- Anne Grosbol Jensen 
- Department of Clinical Biochemistry, AUH

**I** am a *medical* doctor during a PhD

## Simple code 
```{r test}
3 * 3
```

## histogram plot 
```{r Raw}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r Cap and label}
#| fig-cap: "Distribution of BMI"
#| label: fig-BMI-histogram
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```


@fig-BMI-histogram


## Discrete plots
```{r geom bar}
#| fig-cap: "group"
#| label: fig-group-geombar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```
According to @fig-group-geombar the groups... 

```{r}
#| fig-cap: "glykemi"
#| label: fig-glykemi-geombar
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```
According to the @fig-glykemi-geombar most of patients did not have glykemi.


```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Putting two plots side by side 

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

```{r side-by-side}
#| label: fig-BMI-glycemia
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glycimia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```


## Style 
Hvis man bruger style selection ændrer den det så det bliver skrevet rigtigt 

Brug denne så styler den hele dokumentet: styler:::style_active_file()


## Plotting two continuous variable 
```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point()
```


```{r}
#| fig-cap: "scatterplot showing BMI against AUC of plasma glukose"
#| label: fig-scat-BMI-glucose
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_hex()
```

## Data wrangling 
### Selecting data 
```{r}
select(post_meal_data, Age)

select(post_meal_data, id, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))
select(post_meal_data, contains("0"))
```

## Renaming specific columns 
```{
small_post_meal_data |>
  rename(auc_glucose = auc_pg)

small_post_meal_data |>
  rename(auc_glukose = auc_pg, auc_insulin = auc_ins)
```

## Select and rename tasks 
```{r}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("...")) 

post_meal_data |>
  rename(P.Glucose.minus5.OGTT = P.Glucose..5.OGTT)
    
post_meal_data |>
  rename(Insulin.minus5.OGTT.X = Insulin..5.OGTT.X)
         

# Task 3.
post_meal_data |> 
  select(id, BMI, Length, Weight, Age) |>
  rename(Height = Length)
```


## Filtering data by row 
```{r}
small_post_meal_data |> 
  filter(glykemi == 1)

small_post_meal_data |> 
  filter(BMI >=25)

small_post_meal_data |>
  filter(Group == "CTR")
```


```{r}
TRUE & TRUE
```

```{r}
small_post_meal_data |> 
  filter(Group == "CTR" , BMI >= 25)

small_post_meal_data |> 
  filter(Group == "CTR" | BMI >= 25)
```


## Modify or add column
```{r}
tidier_post_meal_data |>
  mutate(insulin_glukose_ratio = auc_insulin/auc_glucose,
         Group = if_else(Group == "CTR", "Control", "First-degres relative"))
```


## Practice with filter and mutate 

```{r, task_1}
post_meal_data |>
  filter(BMI >= 20 & BMI <= 40 & Group == "FDR")
```

```{r, task_2}
post_meal_data |>
  mutate(Percent_body_fat = Fat.mass.DXA/Weight*100) |> 
  filter(Percent_body_fat > 30)
```







