---
title: "Reproducible documents"
author: "Anne G Jensen"
format: html
---

## Basics of R

```{r}
1 + 1
```

## Setup code chunk
```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))
```

## Showing the data raw 

```{r first option}
post_meal_data
```

```{r glimpse option}
glimpse(post_meal_data)
```


?colnames


## Bold and kursiv 
- Anne Grosbol Jensen 
- Department of Clinical Biochemistry, AUH

**I** am a *medical* doctor during a PhD

## Simple code 
```{r test}
3 * 3
```

## histogram plot 
```{r Raw}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r Cap and label}
#| fig-cap: "Distribution of BMI"
#| label: fig-BMI-histogram
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```


@fig-BMI-histogram


## Discrete plots
```{r geom bar}
#| fig-cap: "group"
#| label: fig-group-geombar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```
According to @fig-group-geombar the groups... 

```{r}
#| fig-cap: "glykemi"
#| label: fig-glykemi-geombar
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```
According to the @fig-glykemi-geombar most of patients did not have glykemi.


```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Putting two plots side by side 

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

```{r side-by-side}
#| label: fig-BMI-glycemia
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glycimia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```


## Style 
Hvis man bruger style selection ændrer den det så det bliver skrevet rigtigt 

Brug denne så styler den hele dokumentet: styler:::style_active_file()


## Plotting two continuous variable 
```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point()
```


```{r}
#| fig-cap: "scatterplot showing BMI against AUC of plasma glukose"
#| label: fig-scat-BMI-glucose
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_hex()
```




